### 2013-03-06 Paper/paper.tex
Made changes to introduction

% vim:set textwidth=100:
% vim:set fo+=t:

\documentclass[12pt]{article}

\usepackage{amsmath}
\usepackage[nofiglist,notablist]{endfloat}
\usepackage[usenames,dvipsnames]{color}
\usepackage{color}
\usepackage{authblk}
\usepackage{graphicx}
\usepackage{palatino}
\usepackage[activate={true,nocompatibility},final]{microtype}
\usepackage[super,sort&compress]{natbib}
\pagenumbering{arabic}
\parskip = 0.08in \parindent = 0.0in

% Custom macros for author comments
\newcommand{\Alberto}[1]{\color{ForestGreen}#1\normalcolor }
\newcommand{\Justin}[1]{\color{blue}#1\normalcolor}
\newcommand{\Arijit}[1]{\color{magenta}#1\normalcolor}
\newcommand{\Ken}[1]{\color{red}#1\normalcolor}

\author{Arijit~Roy}
\author{Alberto~Perez}
\author{Ken~A.~Dill}
\author{Justin~L.~MacCallum}
\affil{Laufer Center for Physical and Quantitative Biology\\
    Stony Brook University\\
    Stony Brook, NY 11794-5252.}

\title{Predicting the conformational preferences of proteins using a physics-based free energy
method}

\begin{document}

\maketitle

\begin{abstract}

Calculation of free energy differences is of central importance in the simulation of biochemical
systems. It is particularly difficult to calculate between pairs of macromolecular conformations as
well as a computationally expensive task with existing methods. In this work, confinement approach
is used to calculate absolute free energies of biomolecular systems. This method provides two main
advantages: it does not require a reaction coordinate or transition path and it is fast to compute.
Free energy calculated can be decomposed into a per residue contribution in an approximate way. Per
residue free energy allows us to identify the reason behind conformational preferences in
biomolecules. Through out the article we show its use in different challenging modeling problems. In
particular, we show its use in predicting the conformational preference of chamaleon sequences
(sequences with high sequence identity and different folds). This sequence dependent conformational
preferences and per residue free energy decomposition set the stage for the use of this method in
protein design.

\end{abstract}

\section{Introduction}

Protein structure and conformational changes are of central importance in biochemistry. The
populations of different conformational states are determined by their free energies. The ability to
compute these free energies would allow us to understand and make predictions about a wide variety
of biochemical processes---including the mechanism of enzymes, the function of structural proteins,
binding of proteins to other proteins or DNA, and protein folding.

Computer simulations have proven to be a powerful tool for studying biomolecular systems at atomic
detail [ref]. However, calculating the free energies for conformational changes remains a
challenging problem \cite{Meirovitch2007}, primarily because the timescales for such conformational
transitions may be too long to simulate directly. For example, in protein folding the free energy
landscape is very rough \cite{Dill1997,Dill2008}, which limits conformational sampling and leads to
large uncertainties in the relative populations of different states.

Even with advanced methods like replica exchange molecular dynamics [ref], the molecule may be
become trapped in a local minima. Sometimes this problem may be circumvented by defining a pathway
or reaction coordinate connecting the states of interest. The free energy along this reaction
coordinate can then be determined using methods such as umbrella sampling \cite{Torrie1977},
targeted MD \cite{}, or thermodynamic integration \cite{Tironi1994}. However, accurately sampling
this reaction coordinate may still be difficult and require a large computational effort. In the
case of protein folding, it is often difficult to even devise a meaningful reaction coordinate that
could connect the unfolded and folded states efficiently \cite{}.

The calculation of protein conformational free energy has been successfully attempted by a number of
groups \cite{Meirovitch2007,Ytreberg2006,Zheng2008}. In recent years, methods such as the reference
system method \cite{Ytreberg2006}, deactivated morphing \cite{Park2008}, orthogonal space random
walk \cite{Zheng2008}, and the confinement method \cite{Tyka2006,Cecchini2009} have been used to
calculate free energy differences.

In this work we explore the confinement method\cite{Tyka2006,Cecchini2009}, which relies on the fact
that the free energy difference is a state function and thus is independent of the sampling path.
This allows us to use the alchemical thermodynamic cycle shown in Figure \ref{fig:method}. First,
the two end states are slowly ``frozen'' into a very tight, harmonically restrained ensemble and the
free energies for these two processes are calculated. Next, the free energy difference between the
two restrained ensembles is calculated using normal mode analysis. This allows us to calculate the
free energy difference between the two end states without needing to define a physical path or
reaction coordinate connecting the two states.

The confinement has previously been applied to several small proteins [refs]. Here, we use the power
of GPU computing (using graphic cards to run calculations [cite amber gpu]), which allows us to
extend the method to larger systems while maintaining reasonable computation times. We first
demonstrate that our implementation reproduces the previously reported results for small peptides.
Next, we have applied this method to larger proteins in two different case studies. In the first, we
attempt to pick the most native-like structures from a set of models produced during the Critical
Assessment of Structure Prediction (CASP) \cite{} experiment. In the second case study, we examine
the conformational preferences of a series of chameleon proteins designed by Orban and co-workers
\cite{} that can switch between two completely different folds with only small changes in amino acid
sequence.

\section{Results and Discussion}

\subsection{Our implementation of the confiment method reproduces results from the literature}

As a first step, we performed several control experiments to verify that our implementation of the
confinement method produces results compatible with previous calculations reported in the
literature.

The method has previously been applied to a 16 amino acid residue $\beta$-hairpin from protein G,
known as BHP \cite{Cecchini2009}. We calculated the free energy difference between two different
conformations of the peptide: (1) the native conformation, called bhp1, with a two-stranded
$\beta$-sheet; and (2) a conformation, called bhp3, which has a three-stranded $\beta$-sheet.
Analysis of long (4 $\mu$s) equilibrium simulations \cite{Cecchini2009,Krivov2004}  shows that bhp1
is the more favorable configuration by 1.8 kcal/mol. Using the confinement method, we obtain a value
1.7 kcal/mol, which is in good agreement with the equilibrium simulations and with previous
calculations using the confinement method \cite{Cecchini2009}.

\subsection{The confiment method predicts that experimental structures have lower free energies than
computer-generated models}

Previous applications of the confiment method have focused on relatively short peptides, up to 17
residues in length. In this work we apply the confinement method to larger proteins. As a first
test, we calculated the free energies of models submitted by predictors during the CASP9 experiment
(often referred to as decoys) and compared them with the calculated free energy of the corresponding
experimentally determined structure from the Protein Data Bank (PDB). As
Table~\ref{table:casp_control} shows, in 5 out of 6 cases, the confinement method assigns a lower
free energy to the experimentally determined structure than to any of the decoys---as would be
expected. Although differentiating between experimentally determined structures and computer
generated predictions is not a stringent test of a scoring method---for example, examining the
residue-residue packing can easily distinguish between the two [ref Rosetta Holes]---the results
none the less serve as a useful "sanity check".

\begin{table}
\begin{center}
\caption{The confinement method assigns a more favorable free energy to the experimentally
    determined structure than to computer-generated predictions. For each target, we examined as
    many as five predictions submitted by CASP participants. We report the free energy difference
    between the most favorable decoy and the experimentally determined structure. Positive
    $\Delta\Delta G$ values indicate that the experimental structure is predicted to be more
favorable than any of the decoys.}
\label{table:casp_control}
\begin{tabular}{l l l}\hline
    CASP Target  & PDB Identifier & $\Delta\Delta G_{native \to best decoy}$ (kcal/mol) \\ \hline
     T0531       &    2KJX        &          $11.15 \pm 0.70$ \\ \hline
     T0538       &    2L09        &          $-3.00 \pm 0.47$ \\ \hline
     T0540       &    3MX7        &          $16.94 \pm 0.49$ \\ \hline
     T0559       &    2L01        &          $2.10 \pm 0.24$ \\ \hline
     T0560       &    2L02        &          $18.00 \pm 0.49$ \\ \hline
     T0569       &    2KWY        &          $20.01 \pm 0.69$  \\ \hline
\end{tabular}
\end{center}
\end{table}

\subsection{The confinement method correctly predicts the structural preferences of chameleon
sequences}

In general, proteins with similar sequences tend to have similar structures. This idea is the basis
of comparative modeling and fold recognition in protein structure prediction. There are, however,
examples---often referred to as chameleon sequences---of proteins with similar sequences that have
remarkably different structures. Orban and co-workers have designed a sequence of 56-residues that
is marginally stable in one of two possible folds. By mutating key residues in this sequence they
are able to stabilize one fold or the other (see Figure~\ref{fig:orban}). Sequences that adopt a
mixed alpha/beta structure similar to Protein G are denoted as ``GB'', while sequences that form a
three-helix bundle are denoted as ``GA''. We label the two folds as $3 \alpha$ and $4 \beta +
\alpha$. The GA sequences are observed to fold into the $3 \alpha$ fold and the GB sequences fold
into the $4 \beta + \alpha$ fold. One pair of sequences (GA88/GB88) are 88 percent identical in
sequence and differ in seven positions. Another pair (GA95/GB95) are 95 percent identical and differ
in three positions. The last pair (GA98/GB98) differ only in a single tyrosine to alanine mutation.
The fact such small changes in sequence can lead to such dramatic changes in structure is rather
remarkable. Accurately predicting the structural preferences of these structures presents a serious
challenge for computational methods \cite{Alexander2007}$-$\cite{Shortle20009}.

We initially approached this problem by making a model of each sequence with the same backbone
structure as its partner chameleon sequence. For example, we took the sequence of GA88 and built a
model with the same overall structure as GB88. We then used the confinement method to assess the
free energy difference between the experimentally determined structure of GA88 and the model (with
the GA88 sequence and the GB88 structure). The confinement method was able to predict the
conformational preferences correctly for all six sequences (data not shown). This is, however, not
surprising. It is well known [CITE] that it is easy to distinguish computational models from native
structures. To avoid this potential problem, we instead computed relative free energy of two
different computer generated models for each sequence. One model is based on the GA structure and
the other on the GB structure (see Supporting Information for details on the modeling procedure).
This is a much more realistic test of the confinement method's ability to accurately calculate
relative free energies.

The results of these calculations are presented in Figure~\ref{fig:orban}. The confinement method
identifies the correct structure for all five sequences. One hypothesis for fold switching of
chameleon sequences is that the structural transitions require states with diminished stability
[ref]. It is believed that if the free energy of the native state and the alternative state are
within a range of 5 Kcal/mol, then it is possible for the native state to be destabilized relative
to the alternative state with only small changes in sequence. The stability of the native state can
decrease for a number reason ranging from chemical modification, breaking of disulphide bonds or
mutations---as is the case here. The calculated free energy differences range from around 3.5 to 5.0
kcal/mol, which is consistent with the above hypothesis \cite{He2008}, \cite{{Alexander2009}}.

\subsection{Per-residue free energy calculation can identify the mechanistic detail behind
conformational preferences}

To understand the mechanism behind the conformational preferences of the different protein
sequences, we decomposed the calculated free energy into per-residue contributions as described in
the method section. This helps us to identify important residues that stabilize a particular
conformation. The per residue free energy, $\Delta \Delta G ((4 \beta + \alpha$) - ($3 \alpha))$ is
shown in Figure~\ref{fig:orban_full}(F). In this plot, a negative peak indicates that a residue
favors the $4 \beta + \alpha$ conformation and a positive peak indicates that the residue favors the
$3 \alpha$ structure.

The GA and GB sequences are nearly identical, so there are some common features observed for all
sequences. Roles of some important residues in stabilizing either $4 \beta + \alpha$ or $3 \alpha$
conformation and possible reason for such conformational preferences are discussed in details in
Figure~\ref{fig:orban_full} and Table~\ref{table:orban_perresidue}. The experimental observations
classified the protein into two parts: Amino acids 9-51 are fully structured in both the folds,
where as residues 1-8 and 52-56 are unstructured in $3 \alpha$, but form $\beta$ strand in $4 \beta
+ \alpha$ structure. Most of the amino acid residues in the region 1-9 favor $4 \beta + \alpha$
structure as can be seen from the plot Figure~\ref{fig:orban_full}(F). We can explain most of the
major peaks in this plot. For example,  first major peak appears at residue 7. Here, the hydrophobic
Leu-7 stabilizes the $4 \beta + \alpha$ conformation as its sidechain is oriented towards the
protein hydrophobic core, whereas in $3 \alpha$ structure it exposed to the solvent
(Figure~\ref{fig:orban_full}(A)). Thus, we hypothesise that a mutation of Leu-7 to a hydrophilic
residue would stabilize the $3 \alpha$ fold. The conformational preferences of most of the other
large peaks can be similarly explained by hydrogen bonding, salt bridge formation, and solvent
exposure of hydrophobic residues. They are listed in Figure~\ref{fig:orban_full} and
Table~\ref{table:orban_perresidue}.

We will now discuss the differences between the $3 \alpha$ and $4 \beta + \alpha$ folds due to the
specific differences between the GA and GB sequences. The differences are in positions 20, 30, and
45. GA95 has Leu-20, Ile-30, and Leu-45; GB95 has Ala-20, Phe-30, and Tyr-45. In the $3 \alpha$
fold, residues 20 and 30 are involved in interactions with residue 49. The peak at position 49
stabilizes the GA sequence more than that of GB sequence. In the $3 \alpha$ fold, Ile-49 has
hydrophobic interactions with Leu-20 and Ile-30 with GA sequence, whereas with GB sequence it has
such interactions with smaller Ala-20 (Figure~\ref{fig:orban_full}(I). Also, in the $3 \alpha$ fold
with the GB sequence, Phe-30 being a larger group can not be completely accommodated in the
hydrophobic core (Figure~\ref{fig:orban_full}(I)) and therefore, partially exposed to the protein
surface. There is an interesting difference for residue at position 45 as well. In GB95 sequence and
$4 \beta + \alpha$ fold there is H-bonding between Tyr-45 and Asp-47
(Figure~\ref{fig:orban_full}(K)), which is absent in the $3 \alpha$ conformation. On the other hand
with GA sequence, this H-bond is absent for both the conformation as there is a Leu at position 45
(Figure~\ref{fig:orban_full}(J)).

\begin{table}
\begin{center}
\caption{Analysis of the per-residue free energy decomposition of GA95 and GB95 sequence reveal that certain residues prefer
either $3 \alpha$  or $4 \beta + \alpha$ structure. These observations and the reason behind such conformational preferences are listed in this
table.}
\label{table:orban_perresidue}
\begin{tabular}{| l | l | l |}
\hline
    Information from per                                 & What is happening in           & What is happening in                      \\
    residue plot                                         & $3 \alpha$ conformation        & $4 \beta + \alpha$  conformation          \\ \hline
    Leu-7 favors                                         & hydrophobic residue is         & oriented towards the                      \\
    $4 \beta + \alpha$ (Figure~\ref{fig:orban_full}(A))  & exposed to the solvent         & hydrophobic core                          \\ \hline
    Gln-11 stabilizes $3 \alpha$                   & Gln-11 has H-bond                         & no such H-bond                           \\
    (Figure~\ref{fig:orban_full}(B))           &  with Glu-15                              &                                           \\ \hline
    Glu-14 favors $4 \beta + \alpha$           & hydrophilic residue partially  & completely exposed                                    \\
    (Figure~\ref{fig:orban_full}(E))           & exposed to the solvent         & to the solvent                                        \\ \hline
    Ala-26 favors  $4 \beta + \alpha$          & hydrophobic residue            & oriented towards the                                   \\
     (Figure~\ref{fig:orban_full}(C))          & oriented towards surface       & hydrophobic core                                       \\ \hline
    Ile-49 favors $3 \alpha$                   & hydrophobic interaction        & sidechain exposed                                       \\
    (Figure~\ref{fig:orban_full}(H) and (I))   & inside the protein             & to the solvent                                          \\ \hline
    Ile-49 favors GA more                      & more hydrophobic               &  unchanged                                              \\
    than GB                                    & interactions with residues     &                                                         \\
    (Figure~\ref{fig:orban_full}(H) and (I))  &  at 20, 30 in GA                &                                                         \\ \hline
    Val-54 favors $4 \beta + \alpha$           & hydrophobic residue            & hydrophobic interaction                                \\
    Figure~\ref{fig:orban_full}(D))             & exposed to the solvent        & inside the protein                                      \\ \hline
    Glu-56 favors $3 \alpha$                   & hydrophilic residue partially  & completely exposed                                    \\
    (Figure~\ref{fig:orban_full}(G))            & exposed to the solvent         & to the solvent                                        \\ \hline
    Phe-30 in GB sequence                      & exposed to the solvent         & oriented towards the                                  \\
    stabilizes $4 \beta + \alpha$                  & in protein surface             & hydrophobic core                                      \\ \hline
    Tyr-45 in GB favors                      & No H-bonding                   & H-bond with Asp-47                                    \\
    $4 \beta + \alpha$ (Figure~\ref{fig:orban_full}(K))&                 &                                                       \\ \hline
    Asp-47 in GA sequence                      & H-bond with Lys-50             & No H-bond, No Tyr-45                                  \\
    favors $3 \alpha$                          &                         & in GA sequence                                         \\ \hline
\end{tabular}
\end{center}
\end{table}

\begin{figure}
\includegraphics[width=3.5 in,height=4 in]{orban.pdf}
\label{fig:orban}
\caption{Confinement method correctly predicts the structural preferences of six chameleon
sequences. (A) The six sequences used in this study. (B) Each sequence adopts either a Protein
G-like fold (denoted GB) or a three-helix bundle fold (denoted GA). The relative free energies of
the two folds are reported for each sequence.}
\end{figure}

\begin{figure}
\includegraphics[width=6.4 in,height=5.3 in]{orban_full.pdf}
\label{fig:orban_full}
\caption{Sidechain orientation of some important residues which stabilize/destabilize either of the $4 \beta + \alpha$ or $3 \alpha$ conformation.
The details are in Table~\ref{table:orban_perresidue} and the main text.
A. Leu-7.
B. Gln-11 has H-bond with Glu-15 in $3 \alpha$, which is absent in $4 \beta + \alpha$
C. Ala-26
D. Val-54.
E. Glu-14.
F. plot of free energy difference for GA95 and GB95 sequence. Positive peak stabilize $3 \alpha$ form and negative peak
stabilize $4 \beta + \alpha$.
G. Glu-56 .
H. sidechain orientation of Hydrophic Ile-49, along with Leu-20 and Ile-30 with GA95 sequence
I. sidechain orientation of Hydrophic Ile-49, along with Ala-20 and Phe-30 with GB95 sequence
J. Tyr-45 forms H-bond with Asp-47 with GB95 sequence and $4 \beta + \alpha$ conformation
K. No H-bond possible between Leu-45 and Asp-47 with GB95 sequence and $4 \beta + \alpha$ conformation .}
\end{figure}

\subsection{We have applied the confinement method to structure prediction}

We have applied the confinement method as a ``meta-predictor'' for structure prediction. Here, the
task is to correctly identify the most correct models out of a set of ``decoys'' generated by
different methods during the Critical Assessment of Structure Prediction (CASP) experiment. CASP is
a blind test in which different groups apply methods to predict the 3-dimensional structure of
proteins from their sequences. This is done with a either a three week (for humans) or three day
(for fully automated servers) time limit on each target. Each group is allowed to submit five
possible structures, which they are supposed to rank from best to worst. We have performed two
experiments centered on CASP.

In the first experiment, we examined several targets and tried to rank-order the predictions
produced by other groups. In some cases the structures all came from the same group, in others the
structures were drawn from several different groups. The goal of this experiment is to determine if
the physics-based confinement method can correctly identify more native-like structures as having
low free energies.

Our second experiment was to see if the confinement method can identify structures that are missed
by other meta-prediction servers. Most successful meta-prediction servers are based on the idea of
consensus: if many different prediction methods produce similar results, then that is probably a
correct prediction. This is often a powerful heuristic, but it can miss cases where there is a very
good result that is only predicted by one method. We chose several cases where such structures were
missed and assessed if the confinement method can correctly identify these accurate models.

As is common in the CASP experiment, we assess our results in terms of Global Distance Test Total
Score (GDT-TS) \cite{Zemla2003}, which is a C$\alpha$ based measure of structural accuracy. It can
be understood roughly as the percentage of residues that are correctly positioned in the model
(higher is better, range 0 to 100).

We have chosen the initial models on which to test the methods based on server groups that have
traditionally done well in past CASP events.

\subsubsection{The confinement method can correctly rank different models generated by the same methodology
and distinguish the native structure}

The question that we want to answer here is whether we can rank models that have been generated with
the same methodology. In all cases, we have calculated the free energy difference between the native
structure and the submitted models as well. 
%Identifying the native state from a set of models is a very easy test, that most metapredictors
%already handle correctly, but it is also a first proof of principle. 
First we choose a protein BVU3908 from Bacteroides vulgatus whose PDB id and CASP target code are
2L01 and T0559 respectively. The native structure of this 69 residue protein was solved using NMR.
The best predictor group for this target was "BAKER-ROSETTASERVER". We initially checked similarity
between the models submitted by "BAKER-ROSETTASERVER" and discarded two of them from the analysis on
the basis of being too similar to some of the other models. The GDT\_TS score and rmsd values as
shown in Figure~\ref{fig:T0559} indicate that model 1 was predicted correctly, whereas the order of
model 3 and 5 was wrong. The main difference between model 3 and the rest of the models is that the
orientation of the first alpha helix of model 3 is \Alberto{opposite??}. On the other hand, as shown
in Figure~\ref{fig:T0559} the confine and release method not only can differentiate the native
structure from the submitted models, the ranking also correlates perfectly with the GDT\_TS score.

To further test the method we have taken the example of protein BT2368 from Bacteroides
thetaiotaomicron. The PDB id of this 74 amino acid residue protein is 2L02 and CASP target code is
T0560. We have compared the free energy difference between the native structure and the two of the
five submitted models from the group "Splicer". The remaining three models are discarded as again
they are similar to the rest of the models. The comparison of GDT\_TS score as shown in
Figure~\ref{fig:T0560} and published in the final CASP result is only for models with residues 3 to
66. In order to keep consistency, we have also done our analysis with models and crystal structure
consisting residues 3 to 66. As expected, the native state is identified correctly, and the two
other models are ranked in the correct order in accordance with the GDT\_TS score.

\subsubsection{The confinement method can rank models generated by different methods}

Our next test was between models that are produced with different methodologies. We do this by
looking at some of the best CASP submissions from different research groups. For this purpose we
have chosen the x-ray crystallographic structure of fas apoptosis inhibitory protein molecule whose
pdb id and CASP target codes are 3MX7 and T0540 respectively. This protein contain 90 amino acid
residues with 8 beta strands. We have chosen best models from groups "LTB" and "Mufold" for our
analysis, which are labelled as Model 1 and Model 2 in Figure~\ref{fig:T0540}. As presented in
Figure~\ref{fig:T0540} we found that this method is able to correctly order the models which matches
with the GDT\_TS score.

\subsubsection{Per residue free energy calculation identifies the residues that are responsible for
differences between two conformations}

We used one of the CASP targets (pdb id=2KWY, casp id=T0569) to study the per residue contribution
between this NMR structure and the best CASP9 predicted model (from the 'Mufold' group, GDT\_TS=78).
Visualization upon superimposition of two structures reveals that the difference between the two
structure is at the region consist of residues 48 to 65, where model1 is much more disordered
compare to the native structure(see Figure~\ref{fig:T0569}). Other regions look similar for both the
cases with of model1 is within 2.6 \AA backbone rmsd value of the native structure. We found that,
the confinement method can still differentiate between the native structure and the generated model.
We have also calculated the per residue free energy with a aim that this can identify the residue
which are responsible for such free enerfy difference. The per residue free energy decomposition is
shown in  Figure~\ref{fig:T0569_per_residue}(A). We indeed found that two hydrophobic residues
Val-59 and Ile-61 are oriented towards the solvent in the protein surface in the generated model
(Figure~\ref{fig:T0569_per_residue}(B). This the region, where there is a beta sheet in the original
crystal structure, which is absent in the model. Interestinlgy, some regions are more stable in the
generated model. A close inspection reveals that the hydrophilic Glu in position 67 is completely
exposed to the protein surface and solvent in the best model, whereas it is partially exposed in the
crystal (Figure~\ref{fig:T0569_per_residue}(C)). The peak at position 76 is due to the H-bond
between Lys-76 and Asp-11, which is absent in the original crystalographic structure
(Figure~\ref{fig:T0569_per_residue}(D)).

\subsubsection{Failures in the confinement method}

Despite the great success in most of the studied systems, we found a few failures, specially when
the GDT scores of the compared structures are very close. We have found this for an engineered
protein from Asr4154 protein (PDB ID: 2L09 and CASP Target T0538). This protein contains 54 amino
acid residues with three alpha helixes and two very short beta strands. The model that is closest to
the native structure was generated by the group PconsR with a GDT\_TS value of 96.23. We have also
considered one model from group "Shell" (GDT\_TS = 90.09) and "FOLDIT" (GDT\_TS = 86.32) for our
analysis along with the native crystal struture (RMSD values between 1.6 and 2.0\AA \Alberto{from
native, right?all atom? CA?}). The models from PconsR, Shell and FOLDIT are labelled as model 1,
model 2 and model 3 respectively in Figure~\ref{fig:T0538}. As seen in Figure~\ref{fig:T0538}, one
can find that the model from PconsR has lowest free energy value, making it a more stable structure.
Although ranking of the other models remains consistent, model 1 is predicted to be more stable than
the crystal structure. To rationalize this, we have decomposed the total free energy into the per
residue components. The results show us that despite the small variation at the backbone level (as
shown by high GDT\_TS scores and low RMSDs), the sidechains are oriented in very different ways,
giving raise to large differences in the stabilization of certain residues. In particular, some of
the differences arise from different salt bridge patterns (Arg-32 with Glu-35 and Glu-28 with Lys-24
in crystal versus Arg-32 with Glu-28 in model 1) and certain flexible polar residues exposed to the
surface (Lys-24 in model 1, which has an entropic gain from not forming a salt bridge and is
stabilized by interactions with the solvent). This is summarized in Figure~\ref{fig:T0538compare}(A)
Similar patterns arise in other places of the protein (see Figure ~\ref{fig:T0538compare}(B)), where
the salt bridge between Arg-26 and Glu-50 is absent in model 1 and is further \Alberto{stabilized?}
by Phe-51 having a very different conformation that in the crystal structure. All in all, this
unexpected result just shows us that this method is sensitive to local interactions such as those
happening from side chain reorientation. At the same time, the results explained in past sections
showed that for larger changes the accuracy of the method is very high.

\subsubsection{What can we say about low resolution models?}

So far we have seen that the method is good at predicting preferences when the structures are not
very close to each other, how far from native can we go? In this section we want to explore what
happens when we use the method to describe low quality models. We use one of the CASP targets (pdb
id = 2KJX, CASP id = T0531) . We mentioned earlier that our predictive abilities are greatly
decreased with low model qualities. Comparing low qualitive models between themselves is not very
informative, however, comparing all of them to a native/native like structure, allows to rank order
this structures. In order to test this hypothesis we performed another test with target T0531 of
CASP9, in which we compared the crystal structure to five models from the MUFOLD server. Pdb id of
this extracellular domain of the jumping translocation breakpoint protein is 2KJX and CASP target
code is T0531. Looking at Figure~\ref{fig:T0531} shows: 1.) native is correctly identified as
expected and 2.) Surprisingly there is a high level of correlation between the GDT ranking and the
free energy ranking for model 1 and model 3, the rest three structures with GDT\_TS score less than
35  are ordered incorrectly. It is worth to note that models 2 and 3 have the same GDT and very
different free energies, meaning that the actual ordering could change a lot \cite{Perez2012}. It is
encouraging that at least the method can pick out the best model even though it is got a low GDT
score: 44.

\subsubsection{Can the confinement method perform better quality assessment in protein structure prediction?}

A part of the CASP experiment is dedicated to the assessment of the quality of predicted models
\cite{Kryshtafovych2011}. Most of the top performing groups in this category use consensus
approaches for quality assessment \cite{Wang2011}. A review of CASP9 quality assessment also pointed
out that the methods which are based on clustering techniques perform much better compared to those
methods which are based on analysis of individual models \cite{Kryshtafovych2011}. In this
background we try to answer, whether the confinement method can do better quality assessment
compared to the best performing group (MUFOLD-WQA) in this catagory. In CASP9, the overall
reliability of models in quality assessment (QA) was accepted in a mode known as QMODE 1. Here,
predictors were asked to score each model on a scale from 0 to 1, with higher values corresponding
to better models \cite{Kryshtafovych2011}. We have again picked up the CASP taget T0538 to answer
this question. We have chosen Model 3 submitted by PconsR (GDT\_TS = 96, QMODE 1 = 0.5434) and model
5 from the MULTICOM-NOVEL ( GDT\_TS = 83, QMODE 1 = 0.5865). Both are server predicted models and
they were further used for quality assessment by the group MUFOLD-WQA \cite{Wang2011}. It is
important to note that the model from PconsR was the most accurate predicted model for this target
and the model from MULTICOM-NOVEL was predicted as best by the quality assessment experiments of
MUFOLD-WQA. Also, the QMODE 1 value presented here is from the group MUFOLD-WQA. Our free energy
analysis predict that the model from Pconsr is indeed better than the model from MULTICOM-NOVEL
$(\Delta \Delta G (Model\_Pconsr - Model\_MULTICOM-NOVEL = -3.9 Kcal/Mol)$, which   matches well
with the GDT\_TS score. On the other hand the MUFOLD-WQA predicted the ranking in the reverse order.

We have also carried out a similar type of analysis for the case of Target 560 from the CASP9
experiment. For that we have chosen the model 1 from the group PconsR (GDT\_TS = 94, QMODE 1 =
0.5178), Bilab\-Enable (GDT\_TS = 87, QMODE 1 = 0.5344) and Multicom\_Refine (GDT\_TS = 74, QMODE 1
= 0.4770). If, one wants to rank order with respect to the GDT\_TS values, the model from PconsR is
the best predicted model, but according to the QA assessment by MUFOLD-WQA, the model from
Bilab-Enable was found to be best predicted model. All the QMODE 1 values are from MUFOLD\-WQA. In
CASP 9, the sequence of 74 residues were given to predict the structure and for quality assessment,
whereas during GDT\_TS analysis only residues 3-66 were considered. When we considered the full 74
amino acid residues, the free energy calculated were $(\Delta \Delta G (Model\_Pconsr -
Model\_Bilab\-Enable = 3.3 Kcal/Mol)$ and $(\Delta \Delta G (Model\_Pconsr - Model\_Multicom\_Refine
= -30.0 Kcal/Mol)$ \Alberto{-30 sounds huge!} which support the QA assessment by MUFOLD-WQA. On the
other hand, if we calculate the same free energy with the trimmed version of the protein (with
residues 3-66), the values of the free energy difference become -1.7 and -31.1 Kcal/Mol
respectively, which support the GDT\_TS score. The summary of this whole observation is presented in
Table~\ref{table:560QA}. There is no doubt that the consensus approach can predict the model quality
in a faster manner. But we expect that the confinement method can predict the model quality in a
relatively expensive but much more accurate way.

\begin{table}
\label{table:560QA}
\begin{center}
\begin{tabular}{| l | l | l | l | l |}
\hline
Model            &  $\Delta \Delta G$ (1-74)  & $\Delta \Delta G$ (3-66) &    GDT\_TS    & QMODE 1  \\ \hline
PconsR           &         0.0                &      0.0                 &    94         &  0.5178   \\ \hline
Bilab\-Enable     &        -3.3                &      1.7                 &    87         &  0.5344   \\ \hline
Multicom\_Refine &        30.0                &     31.1                 &    74         &  0.4770   \\ \hline
\end{tabular}
\end{center}
\caption{Comparison of quality assessment by the group MUFOLD\-WQA and the confinement method of the CASP9 target 560 (pdb id: 2L02). 
The GDT\_TS score obtained from CASP9 website, was based on the trimmed model containing residue 3-66, whereas, the QMODE 1 values are from
quality assessment by the group MUFOLD\-WQA and based on residues 1-74.}
\end{table}

\section{Conclusion}

The last 50 years represent our effort to understand what the native state of a protein
    looks like by using theoretical methods. Experimental evidence coming from NMR and x-ray
    cristallography remain so far the only real ways to solve the structure of the native state.
    However, this methods are expensive and slow. Theoretical methods on the other hand are cheaper,
    but despite improvement still have a long way to go. Often this computational pipelines produce
    ensembles of diverse structures and identifying which structure is more native like is still a
    problem. Here we have shown how the application of the confinement method can help identify
    native like states. The high success rate balances with the computational expense. However, with
    GPU technology and the rate at which computers improve, this computational expense is becoming

less of a bottle neck.

\Alberto{I don't agree with some of the points here, but I think I covered what you meant in the
previous paragraph}

In the recent past there have been a number of studies directed towards understanding protein
structure using theoretical tools. Experimentally, it is relatively easy to identify the native
structure from the data obtained from x-ray crystallography or NMR. On the other hand using
theoretical tools it is often difficult to identify the native structure using both physics based
methods and knowledge based method. In order to rank structure from the available data, some groups
use clustering algorithms and the structure of the most populated cluster is picked out as the most
probable structure. However, it may happen that the structure may be trapped and spend most of its
time in a kinetic trap ***kinetic traps in MD,not other methods, local minima***, which may give a
wrong interpretation of the clustering result. On the other hand, some groups also depend on
potential energy functions to pick out their best structure, lacking an entropic component. In this
work we attempted to identify the native/native like protein strcuture from the decoy using free
energy as a scale. We have used the previously developed confinement method, applying it to larger
systems.

We found that in majority of cases, the native structure is always the most stable and we can rank
protein structures upto 100 amino acid residues long. This method works very well if at least one of
the strcuture in the decoy set has high quality. The main advantage of this method is that it does
not require any reaction coordinate. Thus one can calculate the free energy difference between two
very different conformations. An interesting application from this method that we have exploited is
the partition of free energies on a per residue basis. Although, it is an approximate partition, it
gives important insight regarding the conformational preferences of a particular residue. The use of
this decomposition in understanding the effect of a small number of aminoacids in our test chamaleon
sequences allows us to think that in the future, this method may be used for protein design. The
method can be also give wrong result if all the models in the decoy are very poor. It may also be
difficult to rank structures if all the structures are very very close to the native structure. But
we want to ignore this fact as our main aim is to differentiate between the good and the bad
structure. In this article we have extensively tested this method using examples from CASP This
method can be computationally expensive in actual CASP experiment if one want rank hundreds of
submitted models in CASP experiments. However, in real life experiment with a particular target, we
expect this method can give accurate result. The important fact about confinement method is that all
the part of the calculation can be done independently. There is another big bottlenecks in this
calculation. This has to do with the amount of simulations that have to be performed to confine the
ensemble of conformations close to a microstate into a single microstate. We have tackled this issue
by using GPU (Graphical Processing Unit) technology as opposed to the classical CPUs, which gives us
a two order of magnitude boost in computational efficiency. For example, we have mostly carried out
our calculation with Amber program \cite{Case2005} in GPU computer. With a avarage 56 residue we
found that it take only 4 hour to complete 20 ns of the confinement step. So, with available
computer power this method will be easy to use for real problems.
%There is another issue, which is normal mode analysis calculations are Oh(N**3) and require large
%amounts of memory, effectively limiting the size of the systems under study. In most cases
%quasiharmonic analysis.

This method can be further used for calculating free energy due to very large conformational change.
Another application may be the case of ligand binding. It can also be used to distinguish the most
stable structure during protein design. Decompition of total free energy into the per residue
component help identifying residues that give stability to a particular conformation. Another,
important application may be to study the protein folding kinetics, specially to study the phi value
analysis.

less of a bottle neck.
We foresee great potential for this methodology for two reasons: independence from a reaction
coordinate (just need start and end state) and avility to decompose the free energies on a per
residues basis. This last point can be very useful in protein design. Finally, we have presented
data for structure comparison of proteins, but we are currently exploring the possibility of
using this methodology for getting free energies of ligand-protein binding.

\Alberto{Do we mention this in the text? too technical for conclusions:
There is another issue, which is normal mode
analysis calculations are Oh(N**3) and require large amounts of memory, effectively limiting the
size of the systems under study. In most cases quasiharmonic analysis.
}

\section{Method}

The confinement method has been described in details in ref. by Tyka et.al. \cite{Tyka2006} and
Cecchini et. al. \cite{Cecchini2009}. The basic approach of the confinement approach is the same in
both these papers. However, there are some technical differences. Here we briefly describe the
procedure to compute free energy $\Delta G_{AB}$ between two conformations A and B of a protein. The
basic idea to compute the free energy between conformations A and B is to perform a thermodynamic
cycle:

\begin{enumerate}

\item  Minimization of A and B. These minimized conformations (A* and B*)
       are the reference conformation of that state. During this minimization the backbone is kept
       restrained to the original position to avoid large conformational changes.

   \item  The free energy of confining the ensemble (A or B) to a microstate (A* or B*) is
       calculated. This is done by gradually applying larger and larger
       harmonic restraints on all the atoms of the biomolecule. This is done by running 21 molecular dynamics simulation 
       (20 ns long each), where the harmonic restraint force constant was scaled from 0.00005
       \Alberto{units??}
       (mostly free) to 81.92 (frozen in one microstate). In this final restrained state, the
       rotational contribution to the free energy is frozen out
       and the only remaining contribution is the vibrational part. The free energy for this step is
       estimated from the fluctuations around
       the reference structure using a numerical
       approach developed by Tyka et. al. \cite{Tyka2006}. The confinement free energy calculated in
       this way is recorded as 
       $\Delta G_{A,A*}$ and $\Delta G_{B,B*}$ as shown in Figure~\ref{fig:method}.

\item  Finally the thermodynamic cycle is closed by calculating the free energy between the final
       restrained state A* and B* using normal mode analysis or quasiharmonic analysis. The free energy calculated in 
       this way is shown as $\Delta G_{A*,B*}$ in Figure~\ref{fig:method}.

\item  The full free energy, $\Delta G_{A,B}$ between the two state A and B is calculated using the equation 
       $\Delta G_{A,B}$ = $\Delta G_{A,A*}$ - $\Delta G_{B,B*}$ + $\Delta G_{A*,B*}$

\end{enumerate}

All calculations where performed with the amber 11 suit of programs in combination with ff99SB and
GB/SA implicit solvent. Interestingly, we extend the method for calculation of per residue free
energy in an approximate way. For this purpose, the confinement energy, $\Delta G_{A,A*}$ and
$\Delta G_{B,B*}$ of each residue is calculated in the usual numerical way as described above. The
internal energy of each residue is calculated using the decomp module of amber from the final
restrained trajectory. We call this method approximate as we ignore the contribution from the normal
mode or quasiharmonic analysis. However, this contrbution to the total free energy is much smaller
(less than $1.0 Kcal/Mol$\Alberto{that sounds huge, i would say what percent of error we stimate, I
think it was lower than 1kcal/mol}), which allow us to study the mechanistic details of
conformational preference of each residue.

%There are two big bottlenecks in this calculation. The first has to do with the amount of
%simulations that have to be performed to confine the ensemble of conformations close to a
%microstate into a single microstate, the second bottleneck has to do with the size of the system in
%the normal mode calculation. We have tackled the first issue by using GPU (Graphical Processing
%Unit) technology as opposed to the classical CPUs, which gives us a two order of magnitude boost in
%computational efficiency. Normal mode analysis calculations are Oh(N**3) and require large amounts
%of memory, effectively limiting the size of the systems under study. Where as this can be overcome
%by larger memory machines, these are often not available, and so we propose an alternative by using
%quasiharmonic analysis. This is still an Oh(N**3) problem, but we now the motions of atoms are
%already determined in trajectories, and so we can choose to do quasiharmonic analysis excluding
%hydrogens, which will reduce the number of atoms N to almost N/2. This means that we can use this
%method for systems twice as large.

\begin{figure}
\begin{center}
\includegraphics[width=3.5 in,height=3.5 in]{method.pdf}
\end{center}
\caption{Graphical representation of the thermodynamic cycle involving confinement method.}
\label{fig:method}
\end{figure}

%\begin{figure}
%\begin{center}
%\includegraphics[width=3.5 in,height=3.0 in]{G95.pdf}
%\end{center}
%\caption{Two protein with 95 \% similar sequence but different folds}
%\label{fig:G95}
%\end{figure}

\begin{figure}
\begin{center}
\includegraphics[width=3.8 in,height=3.0 in]{T0559.pdf}
\end{center}
\caption{The native and three submitted model structure, along with their GDT\_TS, RMSD and relative Free energy values 
of protein BVU3908 from Bacteroides vulgatus (PDB id: 2L01 and CASP code: T0559).}
\label{fig:T0559}
\end{figure}

\begin{figure}
\begin{center}
\includegraphics[width=3.8 in,height=3.0 in]{T0538.pdf}
\end{center}
\caption{The native and three model structure of engineered protein from Asr4154 protein (PDB ID: 2L09 and CASP code:T0538). The model 1,2 and 3 are
from the group PconsR, Shell and FOLDIT respectively.}
\label{fig:T0538}
\end{figure}

\begin{figure}
\begin{center}
\includegraphics[width=4.9 in,height=4.0 in]{Target_538_compare.pdf}
\end{center}
\caption{Difference of the sidechain orientation between model 1 and crystallographic structure in Target-538.
A. The salt bridge between Glu-35 and Arg-32 in crystal structure. In model 1, this is compensated by H-bonding between 
Arg-32 and Glu-28. There is another salt bridge between Glu-28 and Lys-24.
B. Salt bridge between Arg-26 and Glu-50 in crystal structure. Orientation of Phe-51 is different in crystal than in model1.}
\label{fig:T0538compare}
\end{figure}

\begin{figure}
\begin{center}
\includegraphics[width=3.5 in,height=3.0 in]{T0560.pdf}
\end{center}
\caption{Native and two model structure of protein BT2368 from Bacteroides thetaiotaomicron (pdb id: 2L02 and CASP code: T0560). The two models
were from the group "Splicer".}
\label{fig:T0560}
\end{figure}

\begin{figure}
\begin{center}
\includegraphics[width=3.5 in,height=3.0 in]{T0540.pdf}
\end{center}
\caption{X-ray crystallographic structure and two submitted models of fas apoptosis inhibitory protein (pdb id: 3MX7 and CASP code: T0540). 
Model 1 and Model 2 in this analysis were submitted by the group LTB and MUFOLD respectively.}
\label{fig:T0540}
\end{figure}

\begin{figure}
\begin{center}
\includegraphics[width=3.5 in,height=5.7 in]{T0531.pdf}
\end{center}
\caption{The native structure and 5 models of extracellular domain of the jumping translocation
breakpoint protein (pdb id: 2KJX and the CASP code: T0531).}
\label{fig:T0531}
\end{figure}

\begin{figure}
\begin{center}
\includegraphics[width=3.2 in,height=3.8 in]{T0569.pdf}
\end{center}
\caption{Native and best model structure of a domain of adhesion exoprotein from Pediococcus pentosaceus (pdb id: 2KWY and CASP code: T0569).
The best model was from the group "Mufold".}
\label{fig:T0569}
\end{figure}

\begin{figure}
\begin{center}
\includegraphics[width=5 in,height=4 in]{Target_569_per_residue.pdf}
\end{center}
\caption{A. Plot of per residue free energy difference between the model and the crystal structure of Traget 569 (pdb id: 2KWY) of CASP9.
B. In model 1, the beta sheet containing Val-59 and Ile-61 is disordered. These hydrophobic residues are exposed to the solvent.
C. Glu-67 is partially exposed to the solvent in crystal and it is fully exposed in model 1.
D. There is a H-bond between Lys-76 and Asp-11 in model 1, which is absent in the crystallaographic structure.}
\label{fig:T0569_per_residue}
\end{figure}

\begin{thebibliography}{99}

\bibitem{Meirovitch2007}
Meirovitch, H. Recent developments in methodologies for calculating the entropy and free energy of biological systems by computer simulation.
Current Opinion in Structural Biology, 2007, 17, 181-186.

\bibitem{Chipot2007}
Chipot, C.; Shell, M.S.; Pohorille, A. Introduction, in Chipot, C., Pohorille, A., editors. Free Energy
Calculations: Theory and Applications in Chemistry and Biology. Springer Series in Chemical
Physics, vol. 86. Berlin and Heidelberg: Springer; 2007, p. 1–32.

\bibitem{Jorgensen2004}
Jorgensen, W.L. The many roles of computation in drug discovery, Science 2004, 303, 1813–8.

\bibitem{Gilson2007}
Gilson, M.K.; Zhou, H.X. Calculation of protein-ligand binding affinities. Annu Rev Biophys Biomol Struct. (2007) 36, 21-42.

\bibitem{Torrie1977}
Torrie, G. M.; Valleau, J. P. iNonphysical sampling distributions in Monte Carlo free-energy estimation: Umbrella sampling 
(1977) J. Comput. Phys. 23, 187

\bibitem{Tironi1994}
Tironi, I.G.; van Gunsteren, W.F. A molecular-dynamics simulation study of chloroform. Mol. Phys. (1994) 83, 381-403.

\bibitem{Dill1997}
Dill, K.A.; H.S. Chan.  From Levinthal to Pathways to Funnels:  The "New View" of Protein Folding Kinetics.  Nature Structural Biology 4, 10-19 (1997)

\bibitem{Dill2008}
Dill, K.A.; Ozkan, S.B.; Shell, M.S.; Weikl, T.R. The protein folding problem. Annual Review of Biophysics (2008), 37, 289-316.

\bibitem{Anfinsen1973}
Anfinsen. C.B. Principles that Govern the Folding of Protein Chains. Science (1973) 181, 223-230.

\bibitem{Christ2007}
Christ, C.D.; van Gunsteren, W.F. Enveloping distribution sampling: A method to calculate free energy differences from a single simulation,
J. Chem. Phys. (2007), 126, 184110.

\bibitem{Ytreberg2006}
Ytreberg, F.; Zuckerman, D. Simple estimation of absolute free energies for biomolecules. J. Chem. Phys. 2006, 124, 104105.

\bibitem{Park2008}
Park, S.; Lau, A.; Roux, B. Computing conformational free energy by deactivated morphing. J. Chem. Phys. 2008, 129, 134102

\bibitem{Zheng2008}
Zheng, L.; Chen, M.; Yang, W. Random walk in orthogonal space to achieve efficient free-energy simulation of complex systems, Proc. Natl. Acad. Sci. 2008, 105 (51), 20227.

\bibitem{Tyka2006}
Tyka, M.; Clarke, A.; Sessions, R. An Efficient, Path-Independent Method for Free-Energy Calculations. J.Phys.Chem. B 2006, 110, 17212-17220.

\bibitem{Cecchini2009}
Cecchini, M., Krivov, S.V., Spichty, M., Karplus, M. Calculation of free-energy differences by confinement simulations. Application to peptide conformers. 
J. Phys. Chem. B 113, p. 9728-9740 (2009).

\bibitem{Strajbl2000}
Strajbl, M.; Sham, Y.Y.; Villà, J.; Chu, Z.-T.; Warshel, A. Calculations of Activation Entropies of Chemical Reactions 
in Solution. (2000) 104, 4578-4584.

\bibitem{Krivov2004} 
Krivov, S.; Karplus, M. Hidden complexity of free energy surfaces for peptide (protein) folding Proc. Natl. Acad. Sci. U.S.A. 2004, 101, (41), 14766.

\bibitem{Alexander2007}
Alexander, P.A.; He, Y.; Chen, Y.; Orban, J. Bryan, P. The design and characterization of two proteins with $88 \%$ sequence identity but different 
structure and function. Proc. Natl. Acad. Sci. 2007, 104 (29), 11963-11968.

\bibitem{He2008}
He, Y.; Chen, Y.; Alexander, P.A.; Orban, J. NMR structures of two designed proteins with high sequence identity but different fold and function. Proc. Natl. Acad. Sci. 2008, 105 (38), 14412-14417.

\bibitem{Alexander2009}
Alexander, P.A.; He, Y.; Chen, Y.; Orban, J. Bryan, P. A minimal sequence code for switching protein structure and function. Proc. Natl. Acad. Sci. 2009, 106(50), 21149-21154.

\bibitem{Shortle20009}
Shortle, D. One sequence plus one mutation equals two folds. Proc. Natl. Acad. Sci. 2009, 106(50), 21011-21012.

\bibitem{Sheffler2009}
Sheffler, W.; Baker, D. RosettaHoles: Rapid assessment of protein core packing for structure prediction, refinement, design, and validation. Protein Science. 2009, 18(1), 229-239.

\bibitem{MacCallum2011}
MacCallum, J.; Perez, A.; Schnieders, MJ.; Hua, L.; Jacobson, M.P.; Dill, K.A. Assessment of protein structure refinement 
in CASP9. Proteins, 2011, 79, 74-90.

\bibitem{Kryshtafovych2011}
Kryshtafovych, A.; Fidelis, K; and Tramontano, A. Evaluation of model quality predictions in CASP9. Proteins, 2011, 79, 91–106

\bibitem{Zemla2003}
Zemla, A. LGA: a method for finding 3D similarities in protein structures. Nucleic Acids Res 2003, 31, 3370–3374.

\bibitem{Perez2012}
Perez, A.; Yang, Z.; Bahar, I.; Dill, K.A.; MacCallum, J.L.; FlexE: Using Elastic Network Models to Compare Models of Protein Structure. J. Chem. Theory Comput., 2012, 8, 3985-3991.

\bibitem{Wang2011}
Wang, Q.; Vantasin, K.; Xu, D.; Shang, Y. MUFOLD-WQA: A new selective consensus method for quality assessment in protein structure prediction. 
Proteins, 2011, 79: 185–195.

\bibitem{Case2005}
Case, D.A.; Cheatham, III, T.E.; Darden, T.; Gohlke, Luo, H.R.; Merz, Jr., K.M.;  Onufriev, A; Simmerling, C.; 
Wang, B.; R. Woods, R. The Amber biomolecular simulation programs. J. Computat. Chem. (20005) 26, 1668-1688.

\end{thebibliography}

\end{document}