### 2013-05-03 Paper/corrected.tex

% vim:set textwidth=100:
% vim:set fo+=t:

\documentclass[12pt]{article}

\usepackage{amsmath}
\usepackage[nofiglist,notablist]{endfloat}
\usepackage[usenames,dvipsnames]{color}
\usepackage{color}
\usepackage{authblk}
\usepackage{graphicx}
\usepackage{palatino}
\usepackage[activate={true,nocompatibility},final]{microtype}
\usepackage[super,sort&compress]{natbib}
\pagenumbering{arabic}
\parskip = 0.08in \parindent = 0.0in

% Custom macros for author comments
\newcommand{\Alberto}[1]{\color{ForestGreen}#1\normalcolor }
\newcommand{\Justin}[1]{\color{blue}#1\normalcolor}
\newcommand{\Arijit}[1]{\color{magenta}#1\normalcolor}
\newcommand{\Ken}[1]{\color{red}#1\normalcolor}

\author{Arijit~Roy}
\author{Alberto~Perez}
\author{Ken~A.~Dill}
\author{Justin~L.~MacCallum}
\affil{Laufer Center for Physical and Quantitative Biology\\
    and Departments of Physics and Chemistry\\
    Stony Brook University\\
    Stony Brook, NY 11794-5252.}

\title{Predicting the conformational preferences of proteins using a physics-based free energy
method}

\begin{document}

\maketitle

\begin{abstract}

Calculation of free energy differences is of central importance in the simulation of biochemical
systems. It is particularly difficult to calculate between pairs of macromolecular conformations as
well as a computationally expensive task with existing methods. In this work, confinement approach
is used to calculate absolute free energies of biomolecular systems. This method provides two main
advantages: it does not require a reaction coordinate or transition path and it is fast to compute.
Free energy calculated can be decomposed into a per residue contribution in an approximate way. Per
residue free energy allows us to identify the reason behind conformational preferences in
biomolecules. Through out the article we show its use in different challenging modeling problems. In
particular, we show its use in predicting the conformational preference of chamaleon sequences
(sequences with high sequence identity and different folds). This sequence dependent conformational
preferences and per residue free energy decomposition set the stage for the use of this method in
protein design.

\end{abstract}

\section*{Introduction}

In some problems of protein science, you want to know the relative stability of a protein's conformation $A$ compared to
its conformation $B$.  We call this the \emph{difference free energy}. For example, in all allosteric mechanisms, a protein
often adopts one conformation when a ligand is bound and another conformation when no ligand is bound~\cite{Elber2007}.
To understand the mechanism requires knowledge of the relative free energies of the protein conformations.  Another
example is in protein structure prediction, for example as practiced in the Critical Assessment of Structure Prediction
(CASP) competition~\cite{Moult2011}. If you are using some computational model and have predicted two putative native
structures, $A$ and $B$, you want to compute which has the lower free energy, in order to know which is the more
native-like.  In a third example, you may want to know whether mutating a few amino acids in a protein could cause the
protein to switch from one stable conformation to another, because that could have important consequences for biological
mechanisms and disease. And fourth, sometimes binding a ligand induces a protein from one conformation to another. A
quantitative understanding of such \emph{induced-fit} situations requires knowledge of the difference free energy of the
protein in the two states. In all these cases, it would be useful to have a computational method that can be used
efficiently with atomically detailed physical forcefield models to compute the free energy differences between different
given protein conformations.

One widely explored strategy
is to use molecular dynamics simulations along some putative reaction coordinate pathway from conformation $A$ to
$B$~\cite{Elber2007,West2007,Haas2007,Jonsson1998,E2007,Dellago2002,Cheng2006,Elber2005,Chipot2007}. The free energy along this
reaction coordinate can then be determined using methods such as umbrella sampling \cite{Torrie1977,Mascarenhas2013}  
and weighted histogram analysis method (WHAM) \cite{Kumar1992}. Such
approaches have several limitations. First, it is necessary to know an efficient reaction pathway from $A$ to $B$. If
conformations $A$ and $B$ are quite different, then it can be challenging to find such paths. Second, these methods are
computationally slow. To get an accurate estimate of the total free energy difference $\Delta G = G_B - G_A$ requires
accurate determinations of the many small free energy differences $A \rightarrow 1 \rightarrow 2 \ldots \rightarrow B$,
so each step requires substantial amounts of sampling.  And third, these methods are prone to large errors, because the
pathway error is a sum of many errors among the many steps.

The calculation of protein conformational free energy has been successfully attempted by a number of groups
\cite{Shell2010,Ytreberg2006,Zheng2008,Spichty2010,Strajbl2000,Park2008,Tyka2006,Cecchini2009,Ovchinnikov2013}. Some of
these methods like the reference system method~\cite{Ytreberg2006}, deactivated morphing~\cite{Park2008}, the
confinement method~\cite{Tyka2006,Cecchini2009,Ovchinnikov2013} take an alternative strategy for computing difference
free energies that does not require knowing a pathway from $A$ to $B$.

Here, we adopt the confinement method of Tyka et al~\cite{Tyka2006} and Cecchini et al.~\cite{Cecchini2009}, which is
based on the thermodynamic cycle shown in Figure~\ref{fig:method}. We start with an ensemble of related conformations,
$A$. We then restrict $A$ to a much ``tighter'' ensemble $A^\ast$ by applying position restraints in a series of MD
simulations. Similarly, a second ensemble $B$ is restricted to $B^\ast$. We then compute the free energy between
$A^\ast$ and $B^\ast$ using either normal mode analysis \cite{Brooks1983,Case1994} or the quasiharmonic 
method \cite{Karplus1981,Levy1984}. This step takes into account
both the remaining conformational entropy in each ensemble and the remaining enthalpy difference. In this way, we can
calculate the free energy difference between the two end states without needing to define a physical path or reaction
coordinate connecting the two states. The confinement method shares some similarities with the `confine-and-release'
method for computing ligand binding affinities~\cite{Mobley2007,Mobley2012,Mobley2006}.

\begin{figure}
    \begin{center}
        \includegraphics[width=3.5 in]{method.pdf}
    \end{center}
    \caption{Graphical representation of the thermodynamic cycle involving confinement method.}
\label{fig:method}
\end{figure}

Previously, the confinement method has been validated on small model peptides~\cite{Tyka2006,Cecchini2009}. Here, we
show that the confinement method can also be applied successfully to larger systems, such as the series of chameleon
proteins designed by Orban and co-workers~\cite{Alexander2007,He2008,Alexander2009,Bryan2010,He2012}, which can switch between two
completely different folds with only slight changes in sequence. We also show that our computed difference free energies
are useful in evaluating the quality of CASP target protein predictions.  This could be of significant value for
ultimately improving the energetics in protein-structure models, not just the structures. And, finally, we show that we
can approximately decompose difference free energies into individual amino acid components. This offers the opportunity
for diagnosing the structural basis for difference free energies, which can be useful for interpreting biological
mechanisms.

\section*{Results and Discussion}

\subsection*{The confinement method succeeds at some basic consistency checks}

First, we validated that our implementation of the confinement method produces results compatible with previous
calculations reported in the literature.  The method has previously been applied to a 16 amino acid residue
$\beta$-hairpin from protein G, known as BHP~\cite{Cecchini2009}. We calculated the free energy difference between the
native conformation---called bhp1, which has a two-stranded $\beta$-sheet---and a non-native conformation---called bhp3,
which has a three-stranded $\beta$-sheet. Our confinement calculation shows that bhp1 is more stable by 1.7 kcal/mol,
consistent with 4 $\mu$s equilibrium simulations showing that bhp1 is more favorable configuration by 1.8 kcal/mol, and
in agreement with previous calculations~\cite{Cecchini2009}.

Second, we looked at 6 target proteins from the CASP9 experiment~\cite{Moult2011}.  For each target, we examined up to 5 submitted
models.  We computed the difference free energy between the true native and the best model.
Figure~\ref{fig:summary_casp} and Table~\ref{table:casp_control} show that, in 5 out of 6 cases, the confinement method
assigns a lower free energy to the experimentally determined structure than to any of the decoys.  Other well-known
discriminators can also successfully tell native structures from computer generated models~\cite{Sheffler2009,Zhou2002} 
; it is just an independent useful validation that the confinement calculations make
sense.

\begin{table}
\begin{center}
\caption{The confinement method assigns a more favorable free energy to the experimentally determined structure than to
    computer-generated predictions. For each target, we examined as many as five predictions submitted by CASP
    participants. We report the free energy difference between the most favorable decoy and the experimentally
    determined structure. Positive $\Delta\Delta G (=G_{best~decoy} - G_{native})$ values indicate that the experimental structure is predicted to be
    more favorable than any of the decoys.}
\label{table:casp_control}
\begin{tabular}{l l l l}\hline
    CASP Target  & PDB Identifier & $\Delta \Delta G$ (kcal/mol) & GDT-TS \\ \hline
     T0531       &    2KJX        &          $11.15 \pm 0.70$    &  $44$ \\ \hline
     T0538       &    2L09        &          $-3.00 \pm 0.47$    &  $96$ \\ \hline
     T0540       &    3MX7        &          $16.94 \pm 0.49$    &  $70$ \\ \hline
     T0559       &    2L01        &          $2.10 \pm 0.24$     &  $94$ \\ \hline
     T0560       &    2L02        &          $18.00 \pm 0.49$    &  $94$ \\ \hline
     T0569       &    2KYW        &          $20.01 \pm 0.69$    &  $78$ \\ \hline
\end{tabular}
\end{center}
\end{table}

\subsection*{The confinement method correctly predicts the structures of chameleon sequences}

We also tested the confinement method predictions for difference free energies on the chameleon sequences of Orban et
al.\cite{Alexander2007,He2008,Alexander2009,Bryan2010,He2012}. These are instances in which two highly
similar sequences fold into remarkably different structures. Orban and co-workers have designed a protein-G-like
sequence of 56-residues that is marginally stable in one of two possible folds. By mutating key residues in this
sequence they are able to stabilize one fold or the other (see Figure~\ref{fig:orban}). We refer to the $4\beta +
\alpha$ structure as the $\beta$ conformation, and we refer to the $3\alpha$ structure as the $\alpha$ conformation. One
pair of sequences (GA88/GB88) is 88 percent identical in sequence, differing in seven positions. Another pair
(GA95/GB95) is 95 percent identical, differing in three positions. Accurately predicting the structural preferences of
these structures presents a serious challenge for computational
methods \cite{Allison2011}.

\begin{figure}
    \begin{center}
    \includegraphics[width=5.0in]{orban.pdf}
    \end{center}
    \caption{The confinement method correctly predicts the structural preferences of four chameleon sequences. The top
        part of the figure represent four sequences used in this study along with the protein data bank identifier. The
        experimentally observed fold, computationally predicted fold, and difference free energy between the two folds
        is reported for each sequence.}
\label{fig:orban}
\end{figure}

We initially approached this problem by making a model of each sequence with the same backbone structure as its partner
chameleon sequence. For example, we took the sequence of GA88 and built a model with the same overall structure as GB88.
We then used the confinement method to assess the free energy difference between the experimentally determined structure
of GA88 and the model (with the GA88 sequence and the GB88 structure). The confinement method was able to predict the
conformational preferences correctly for all four sequences (data not shown). This is, however, not surprising as it is
often easy to distinguish computational models from native structures~\cite{Handl2009,Sheffler2009}. To avoid this
potential problem, we instead computed relative free energy of two different computer generated models for each
sequence. One model is based on the $\alpha$ structure and the other on the $\beta$ structure (see Supporting
Information for details on the modeling procedure). This is a much more realistic test of the confinement method's
ability to accurately calculate the difference free energy.

Figure~\ref{fig:orban} shows that the confinement method identifies the correct structure for all four sequences. One
hypothesis is that such structural transitions require states with diminished stability~\cite{Bryan2010}. It is believed
that if the free energy of the native state and the alternative state are within a range of around 5 kcal/mol, then it
is possible for the native state to be destabilized relative to the alternative state with only small changes in
sequence~\cite{He2008,Alexander2009,Bryan2010}. The calculated free energy differences range from around 3.5 to 5.0
kcal/mol, which is consistent with this hypothesis. In a more recent study\cite{He2012}, the amino acid residue at
position 45 (Tyr for $\beta$ and Leu for $\alpha$) was found to be important for switching between $\alpha$ and $\beta$
conformation. This inspired us to introduce another mutation at this position, Y45A, which we refer to as GA98. Our
calculations predict that this mutation shifts the equilibrium to the $\alpha$ conformation, which is now more stable
than the $\beta$ conformation by 3.8 kcal/mol. Although this result has not yet been confirmed experimentally, it is
consistent with the previously observed effect of Y45L.

\subsection*{Per-residue free energy calculations can identify the mechanistic detail behind conformational preferences}

To better understand the mechanism behind these chameleon proteins, we decomposed the calculated free energy into
per-residue contributions in an approximate way~\cite{Tyka2006}. We decompose each confinement step ($\Delta
G_{A,A^\ast}$, $\Delta G_{B,B^\ast}$) into per-residue contributions. We can also decompose the remaining enthalpy in
the confined ensemble into per-residue contributions. However, the method is approximate because we do not include the
residual conformational entropy from the normal mode or quasiharmonic analysis steps. Computing per-residue
contributions helps us to identify important residues that stabilize a particular conformation. The per residue free
energy, $\Delta \Delta G (\beta - \alpha)$ is shown in Figure~\ref{fig:perresidue_orban}.

\begin{figure}
    \begin{center}
        \includegraphics[width=4.5in]{delta_g.pdf}
    \end{center}
    \caption{Per-residue difference free energies for the $\alpha$ and $\beta$ conformations of the GA95 and GB95
        sequences. Residues colored in blue favor the $\alpha$ structure, residues in red favor the $\beta$ structure,
        and white residues have no preference.}
\label{fig:perresidue_orban}
\end{figure}

Although the overall free energy difference between the two structures is small (within 3--5 kcal/mol), individual
residues can show marked preferences for being in either $\alpha$ or $\beta$ conformation. Such differences can be
understood by looking in detail at the local environment for those residues. For example, the region around residue 7
forms a random coil in the $\alpha$ structure, whereas it is forming a beta sheet structure in the $\beta$ structure.
These residues strongly favour the locally well packed and hydrogen bonded environment found in the $\beta$ sheet.
Overall, favoring $\alpha$ or $\beta$ structure is a delicate balance, where the relative global free energy difference
is small and the contributions of different per residue tendencies balance out. It is therefore very likely that by
changing key residue the global fold preferences can be changed.

Three residues are mutated between GA95 and GB95, at positions 20, 30, and 45. In GA95, these residues (L20, I30 and
L45) stabilize the $\alpha$ structure (compare the upper and lower panels on Figure~\ref{fig:perresidue_orban}). In
GB95, two of these residues (F30 and Y45) favor the $\beta$ structure, because they have large solvent exposed surface
areas in the $\alpha$ structure, but are more buried in the $\beta$ structure. Additionally Y45 forms a hydrogen bond
with D47 in the $\beta$ structure. On the other hand, residue A20 from GB95 still favors the $\alpha$ structure,
although less strongly than L20 in GA95.

The $\alpha$ and $\beta$ sequences are nearly identical, so there are some common features observed for all sequences.
The experimental observations classified the protein into two parts: Amino acids 9--51 are fully structured in both
folds, whereas residues 1--8 and 52--56 are unstructured in the $\alpha$ fold, but form $\beta$-strands in the $\beta$
fold. Most of the amino acid residues in the region 1--9 have negative per-residue free energies, which means that these
residues favor the $\beta$ structure. Roles of some other important residues in stabilizing either the $\alpha$ or
$\beta$ conformation are summarized in Figure S1.

In addition to the direct effects of the mutation, there are also indirect effects due to small perturbations in the
environment around the mutations. For example, the L20A mutation causes a slight repacking around residue 20. This
causes large changes in the per-residue free energies of nearby residues T25 and A26. However, the changes for these two
residues have opposite signs and nearly cancel.

These per-residue free energy decompositions provide a great deal of insight into the driving forces behind protein
folding and conformational change. We believe that such calculations may also be useful for protein design---designing
proteins with specific structures and functions.

\subsection*{The confinement method is a useful tool for structure prediction}

We have tested the ability of the confinement method to act as a ``meta-predictor'' for structure prediction. Here, the
task is to correctly identify the most accurate models out of a set of ``decoys'' generated by different methods during
the CASP experiment. CASP is a blind test in which different groups apply methods to predict the 3-dimensional structure
of proteins from their sequences. Each group is allowed to submit five possible structures, which they are supposed to
rank from best to worst.

We have performed two experiments centered on CASP. In the first experiment, we tried to rank-order predictions for
several targets. For each target, the predictions were either produced by a single group---presumably using the same
method for each prediction, or were produced by several different groups---using different methods. The goal of this
experiment is to determine if the physics-based confinement method can correctly identify more native-like structures as
having lower free energies. Our second experiment was to see if the confinement method can identify structures that are
missed by other meta-prediction servers. Most successful meta-prediction servers are based on the idea of consensus: if
many different prediction methods produce similar results, then that is probably a correct prediction \cite{Kryshtafovych2011,Wang2011}. This is
often a powerful heuristic, but it can miss cases where there is a very good result that is only predicted by one
method. We chose several cases where such structures were missed by the best meta-predictors in CASP and assessed if the
confinement method can correctly identify these accurate models.

As is common in the CASP experiment, we assess our results in terms of Global Distance Test Total Score (GDT-TS)
\cite{Zemla2003}, which is a C$\alpha$ based measure of structural accuracy. It can be understood roughly as the
percentage of residues that are correctly positioned in the model (range 0 to 100, higher is better). We did have enough
computer model to analyze every model, so the initial models for the test were chosen from a selection of different
server groups that have done well in past CASP events.

Overall, the confinement method performs well on our CASP tests. Figure~\ref{fig:summary_casp} shows that in almost
every case, the native structure has the lowest free energy and the best model has the next lowest free energy. The
confinement method appears to be a useful tool for ranking structure predictions.

\begin{figure}
\begin{center}
\includegraphics[width=3.5 in]{casp.pdf}
\end{center}
\caption{The confinement method is usually able to identify the native structure and the best model (the model with the
    highest GDT-TS) from a set of decoys.}
\label{fig:summary_casp}
\end{figure}

\subsection*{The confinement method can correctly rank-order structure predictions}

First, we examined the ability of the confinement method to rank different models for a target that have been generated
by a single group using the same methodology for all predictions. We examined two targets: T0559 and T0560 (see Table~\ref{table:casp_testcase}).

\begin{table}
\begin{center}
\caption{The six different CASP targets that are used in this study. The corresponding PDB Identifier and the 
         organism type are also listed.}
\label{table:casp_testcase}
\begin{tabular}{l l l}\hline
    CASP Target  & PDB Identifier &  organism type  \\ \hline
     T0531       &    2KJX        &  extracellular domain of the jumping  \\
                 &                &  translocation breakpoint protein  \\ \hline
     T0538       &    2L09        &  protein asr4154 from Nostoc sp. PCC7120     \\ \hline
     T0540       &    3MX7        &  human Fas apoptotic inhibitory molecule     \\ \hline
     T0559       &    2L01        &  protein BVU3908 from Bacteroides vulgatus   \\ \hline
     T0560       &    2L02        &  protein BT2368 from Bacteroides  \\ 
                 &                &  thetaiotaomicron         \\ \hline
     T0569       &    2KYW        &  domain of adhesion exoprotein from \\ 
                 &                &   Pediococcus pentosaceus \\ \hline
\end{tabular}
\end{center}
\end{table}

The first test case is target T0559. The best predictor group for this 69 amino acid target was ``BAKER-ROSETTASERVER''.
To save computer time, we excluded two models that were very similar to other models that we did include. For this
target, the difference free energy computed by the confinement method can be used to accurately rank-order all of the
models and the native experimental structure (Figure~\ref{fig:T0559}).

\begin{figure}
    \begin{center}
        \includegraphics[width=3.8 in]{T0559.pdf}
    \end{center}
    \caption{The confinement method can correctly rank the native structure and three predictions produced by a single
        method for target T0559.}
\label{fig:T0559}
\end{figure}

We performed a similar calculation for target T0560 with two models from the group ``Splicer''. The remaining three
models were discarded as they were too similar to the rest of the models. Again, we can correctly identify the native
sate and our calculated free energy based ranking matches well with GDT-TS (Figures S2 in the supporting information).

Next, we tested the ability of the confinement method to rank models for target T0540 that were produced by different
prediction groups. The top models from groups ``LTB'' (Model 1) and ``Mufold'' (Model 2) were chosen for analysis. Once
again, the free energy based ranking correlates well with the GDT-TS based score (Figure~\ref{fig:T0540}).

\begin{figure}
    \begin{center}
        \includegraphics[width=3.5 in]{T0540.pdf}
    \end{center}
    \caption{The confinement method can correctly rank the native structure and two models submitted by different
        prediction groups.}
\label{fig:T0540}
\end{figure}

\subsection*{The per-residue free energy is sensitive to small changes in protein conformation:}

In a previous section, we discussed how the per-residue free energy can reveal mechanistic
detail behind the conformational preference of a chameleon sequence. In this section, our aim is to apply the same
method and try to understand if the per-residue free energy can help us identify residues which stabilize or destabilize
a particular region of a protein. In this case, the two conformations have similar folds with only small changes in
localized regions. We chose CASP target T0569 and compared the experimental NMR structure with the best predicted model
(GDT-TS=78; predicted by the ``Mufold'' group). The confinement method predicts that the experimental structure is more stable
by 20 kcal/mol (Figure S2).

The confinement method clearly identifies two hydrophobic residues V59 and I61, which destabilize the predicted model
with respect to the crystallographic structure (Figure~\ref{fig:T0569_per_residue} and Figure S3). The sidechains of
these hydrophobic residues are oriented towards the protein hydrophobic core in the native NMR structure but oriented
towards the exterior of the protein and are solvent exposed in the model. These residues are part of a beta-sheet
in the experimental structure, but because of their sidechain orientation the corresponding beta-sheet become disordered in the predicted model 
(Figure~\ref{fig:T0569_per_residue} and Figure S4). There is also a large
difference around K76, which forms a salt-bridge with D11 in the predicted model, but not in the experimental structure.
This suggests that salt-bridge interactions are too favorable for the combination of force field and implicit solvent
model we use, which has been a problem noted in the past \cite{Roe2007}.

\begin{figure}
    \begin{center}
        \includegraphics[width=3.5in]{T0569_perres1.pdf}
    \end{center}
    \caption{Per-residue difference free energy between the experimental NMR structure and the best prediction for CASP
        target T0569. The amino acid residues that are colored in deep red and deep blue stabilizes the NMR structure
        and the prediction, respectively; the residues with light blue color does not have a strong preference.}
\label{fig:T0569_per_residue}
\end{figure}

\subsection*{The confinement method occasionally produces incorrect results:}

Despite success for most of the studied systems, there are few failures, specially when the GDT-TS scores of the
compared structures are very close. One example is Target T0538, where we compared the experimental structure with three
models (Model 1: ``PconsR''---GDT-TS=96; Model 2: ``Shell''---GDT-TS=90; Model 3: ``FOLDIT''---GDT-TS=86). Contrary to
our expectation, the confinement method predicts that computer generated Model 1 is more stable than the crystal structure (Figure S5).
Per-residue free energy calculations (not shown) show that despite only small variations at the backbone level, the
sidechains are oriented in very different ways (Figure~\ref{fig:T0538compare}), giving rise to large differences in the stabilization of
certain residues. In particular, some of the differences arise from different salt bridge patterns and certain flexible
polar residues exposed to the surface. This unexpected result shows that the confinement method is very sensitive to
local interactions (including sidechain reorientation) and may indicate issues with the forcefield and implicit solvent
models used in the calculation.

\begin{figure}
\begin{center}
\includegraphics[width=4.9 in,height=4.0 in]{Target_538_compare.pdf}
\end{center}
\caption{The per residue free energy calculation found some differences of the sidechain orientation in computer generated model 1 and 
experimental structure of CASP target T0538. 
(A) Two salt bridges between residues E35-R32 and E28-K24 exist in the native structure. These are
absent in the computer generated model structure. Instead
(B) a new salt bridge is formed between residues R32-E28 in model. (C) Another salt bridge between
residues  R26-E50 in the experimental structure. (D) In comparison to the native 
structure, orientation of F51 differs in the model.}
\label{fig:T0538compare}
\end{figure}

\subsection*{Can the confinement method perform better quality assessment in protein structure prediction?}

A part of the CASP experiment is dedicated to the quality assessment (QA) of predicted models \cite{Kryshtafovych2011}.
Here, predictors were asked to score each model on a scale (known as qmode) from 0 to 1, with higher values
corresponding to better models \cite{Kryshtafovych2011}. Most of the groups use consensus strategies in such experiments.
Here, we chose two computer generated models from CASP taget T0538, where the top performing group "MUFOLD-WQA" \cite{Wang2011} failed.
One of the model was from "PconsR" (GDT-TS = 83, qmode 1 = 0.5865) and another from "MULTICOM-NOVEL" ( GDT-TS = 83, qmode 1 = 0.5865). We 
choose these two models as the model by PconsR was the most accurate
predicted model for this target, whereas the other model was predicted best by QA test of MUFOLD-WQA. The calculated
free energy using confinement method indicate that the model by PconsR is more stable by $3.9 Kcal/Mol$ which support
the GDT\_TS trend.  Although consensus prediction strategies do well specially when dealing with large numbers of structures, they 
can miss a good answer that is only predicted by a few methods. At least in this case, the confinement
method is able to capture this structure that was otherwise missed.

\section*{Conclusion}

We have described a computational method called confinement for computing the difference free energy between two
conformational ensembles. We show that the difference free energy can be calculated on ~100 residue sized proteins, even
for large conformational changes. We have demonstrated that it can discriminate the folding preferences of a series of
chameleon proteins. We show that the confinement method can discriminate between the native structure and structure
predictions and can identify the best prediction reliably. Perhaps most importantly, we have shown that it can be used
to give residue-level insights into what are the dominant structural factors in a protein that are responsible for the
difference free energies. The confinement should be useful for protein design, structure prediction, and understanding
the mechanism of conformational change.

A key advantage is that this method does not require any reaction coordinate, or sampling a
pathway from conformation $A$ to $B$.

\section{Method}

The confinement method has been described in details in ref. by Tyka et al. \cite{Tyka2006} and
Cecchini et  al. \cite{Cecchini2009}. The basic approach of the confinement approach is the same in
both these papers. A thermodynamic cycle is used to compute the free energy between conformations A
and B. However, there are some small technical differences between the two approaches. Here we briefly
describe the procedure that we used.

\begin{enumerate}

\item  In the first step, a minimization of A and B conformations are performed. These minimized conformations
    ($A^{\ast}$ and $B^{\ast}$) are the reference conformation of that state.

\item The free energy of confining the ensemble (A or B) to a tighter ensemble ($A^\ast$ or $B^\ast$) is calculated.
    This is done by gradually confining each atom to its position in the reference conformation ($A^{\ast}$ or
    $B^{\ast}$) using a series of progressively stronger position restraints. This is done by running 21 molecular
    dynamics simulation (each 20 ns long) for each leg of the thermodynamic cycle, where the harmonic restraint force
    constant was scaled from 0.00005 Kcal/Mol (mostly free) to 81.92 kcal/mol (tightly restrained). The free energy for
    this step is estimated using a thermodynamic integration approach developed by Tyka et. al. \cite{Tyka2006}. The
    confinement free energies calculated this step are denoted as $\Delta G_{A,A^\ast}$ and $\Delta G_{B,B^\ast}$ in
    Figure~\ref{fig:method}.

\item The thermodynamic cycle is closed by calculating the free energy between the final restrained states $A^\ast$ and
    $B^\ast$ using normal mode analysis \cite{Brooks1983,Case1994} or quasiharmonic analysis \cite{Karplus1981,Levy1984}. The free energy calculated in this way is shown as
    $\Delta G_{A^\ast,B^\ast}$ in Figure~\ref{fig:method}.

\item  The full free energy, $\Delta G_{A,B}$ between the two state A and B is calculated as
    $\Delta G{A,B} = \Delta G{A,A^\ast} - \Delta G{B,B^\ast} + \Delta G{A^\ast,B^\ast}$.

\end{enumerate}

One advantage of confinement method is that
none of the simulation during restraining step depends on each other. Therefore, with available computer resources it can be fast 
to compute. The computational cost for a $56$-residue proteins with chameleon sequence that is reported in this article is only $4$ hours 
in a single gpu computer. All calculations were performed with the Amber 11 suit of programs~\cite{Case2012,Goetz2012} in combination with ff99SB
forcefield~\cite{Hornak2006} and generalized born implicit solvent \cite{Mongan2006}.

Interestingly, we extend the
method for calculation of per residue free energy in an approximate way. For this purpose, the confinement energy,
$\Delta G_{A,A*}$ and $\Delta G_{B,B*}$ of each residue is calculated in the usual numerical way as described in ref. by
Tyka et al. \cite{Tyka2006}. The internal energy of each residue is calculated using the decomp module of amber from the
final restrained trajectory. We call this method approximate as we ignore the entropic contribution from the normal mode
or quasiharmonic analysis. However, this contrbution to the total free energy is much smaller, which allow us to study
the mechanistic details of conformational preference of each residue.

\begin{thebibliography}{99}

\bibitem{Elber2007}
Elber, R. A Milestoning Study of the Kinetics of an Allosteric Transition: Atomically Detailed Simulations of Deoxy Scapharca
Hemoglobin. Biophysical J., 2007, 92, 85-87.

\bibitem{Moult2011}
Moult, J.; Fidelis, K.; Kryshtafovych, A.; Tramontano, A. Critical assessment of methods of protein structure prediction (CASP)-round IX.
Proteins, 2011, 79, 1-5.

\bibitem{West2007}
West, A.M.; Elber, R.; Shalloway, D. Extending molecular dynamics time scales with milestoning: example of complex kinetics
in a solvated peptide. J Chem Phys. 2007, 126, 145104-1451014.

\bibitem{Haas2007}
Haas, K.; Chu, J.W. Decomposition of energy and free energy changes by following the flow of work along reaction path.
J. Chem. Phys. 2009, 131, 144105-144111.

\bibitem{Jonsson1998}
Jonsson, H.; Mills, G.; Jacobsen, K.W. Nudged Elastic Band Method for Finding Minimum Energy Paths of Transitions,
in Classical and Quantum Dynamics in Condensed Phase Simulations, Ed. B. J. Berne, G. Ciccotti and D. F.
Coker, 385 (World Scientific, 1998).

\bibitem{E2007}
E, W.; Ren, W.; Vanden-Eijnden, E. Simplified and improved string method for computing the minimum energy paths in
barrier-crossing eventsi. J. Chem. Phys. 2007, 126, 164103.

\bibitem{Dellago2002}
Dellago, C.; Bolhuis, P.G.; Geissler, P.L. Transition Path Sampling, Adv. Chem. Phys. 2002, 123, 1-84.

\bibitem{Cheng2006}
Cheng, X.; Wang, H.; Grant, B.; Sine, S.M.; McCammon, J.A. Targeted Molecular Dynamics Study of C-Loop Closure
and Channel Gating in Nicotinic Receptors. 2006, 9, 134.

\bibitem{Elber2005}
Elber, R. Long-timescale simulation methods. Cur. Opi. in Str. Biol. 2005, 15, 151-156.

\bibitem{Chipot2007}
Chipot, C.; Shell, M.S.; Pohorille, A. Introduction, in Chipot, C., Pohorille, A., editors. Free Energy
Calculations: Theory and Applications in Chemistry and Biology. Springer Series in Chemical
Physics, vol. 86. Berlin and Heidelberg: Springer; 2007, p. 1–32.

\bibitem{Torrie1977}
Torrie, G. M.; Valleau, J. P. Nonphysical sampling distributions in Monte Carlo free-energy estimation: Umbrella sampling
(1977) J. Comput. Phys. 23, 187

\bibitem{Mascarenhas2013}
Mascarenhas, N. M.; Kastner, J. How maltose influences structural changes to bind to maltose-binding protein: Results from umbrella sampling simulation. Proteins. 2013, 81, 185-198.

\bibitem{Kumar1992}
Kumar, S.; Rosenberg, J.M.; Bouzida, D.; Swendsen, R.H.; Kollman, P.A. The weighted histogram analysis method for free-energy calculations on biomolecules. I. The 
method. J Comput Chem. 1992, 13, 1011-1021.

\bibitem{Ytreberg2006}
Ytreberg, F.; Zuckerman, D. Simple estimation of absolute free energies for biomolecules. J. Chem. Phys. 2006, 124, 104105.

\bibitem{Park2008}
Park, S.; Lau, A.; Roux, B. Computing conformational free energy by deactivated morphing. J. Chem. Phys. 2008, 129, 134102

\bibitem{Zheng2008}
Zheng, L.; Chen, M.; Yang, W. Random walk in orthogonal space to achieve efficient free-energy simulation of complex systems, Proc. Natl. Acad. Sci. 200
8, 105 (51), 20227.

\bibitem{Shell2010}
Shell, S.M. A replica-exchange approach to computing peptide conformational free energies. Mol. Sim. 2010, 7, 505-515.

\bibitem{Tyka2006}Tyka, M.; Clarke, A.; Sessions, R. An Efficient, Path-Independent Method for Free-Energy Calculations. J.Phys.Chem. B 2006, 110, 17212-17220.

\bibitem{Cecchini2009}
Cecchini, M., Krivov, S.V., Spichty, M., Karplus, M. Calculation of free-energy differences by confinement simulations. Application to peptide conformers. J. Phys. Chem. B. 2009, 113, 9728-9740.

\bibitem{Ovchinnikov2013}
Ovchinnikov, V.; Cecchini, M.; Karplus, M. A Simplified Confinement Method for Calculating Absolute Free Energies
and Free Energy and Entropy Differences. J. Phys. Chem. B. 2013, 117, 750-762.

\bibitem{Spichty2010}Spichty, M.; Cecchini, M.; Karplus, M. Conformational Free-Energy Difference of a Miniprotein from Nonequilibrium
Simulations. J. Phys. Chem. Lett., 2010, 1, 1922-1926.

\bibitem{Strajbl2000}
Strajbl, M.; Sham, Y.Y.; Villà, J.; Chu, Z.-T.; Warshel, A. Calculations of Activation Entropies of Chemical Reactions
in Solution. (2000) 104, 4578-4584.

\bibitem{Brooks1983}
Brooks, B. R.; Karplus, M. Harmonic dynamics of proteins: Normal modes and fluctuations in bovine pancreatic trypsin inhibitor. Proc. Natl. Acad. Sci. U.S.A. 1983, 80, 6571-6575.

\bibitem{Case1994}
Case, D. Normal-mode analysis of protein dynamics Curr. Opin. Struct. Biol. 1994, 4, 285-290.

\bibitem{Karplus1981}
Karplus, M.; Kushick, J. Method for estimating the configurational entropy of macromolecules Macromolecules. 1981, 14, 325-332.

\bibitem{Levy1984}
Levy, R.; Karplus, M.; Kushick, J.; Perahia, D. Evaluation of the configurational entropy for proteins: application to molecular dynamics simulations of 
an $\alpha$-helix Macromolecules 1984, 17, 1370-1374.

\bibitem{Mobley2007}
Mobley, D.L.; Chodera, J.D.; Dill, K.A. The combining and release method: obtaining correct binding free energies in
the presence of protein conformational change. Journal of Chemical Theory and Computation 2007, 3, 1231-1235.

\bibitem{Mobley2012}
Mobley, D.L.; Klimovich, P.V. Perspective: Alchemical free energy calculations for drug discovery.
J. Chem. Phys. 2012, 137, 230901-12.

\bibitem{Mobley2006}
Mobley, D.L.; Chodera, J.D.; Dill, K.A. On the use of orientational restraints and symmetry corrections in alchemical
free energy calculations. J. Chem. Phy. 2006, 125, 084902.

\bibitem{Alexander2007}
Alexander, P.A.; He, Y.; Chen, Y.; Orban, J. Bryan, P. The design and characterization of two proteins with $88 \%$ sequence identity but differentstructure and 
function. Proc. Natl. Acad. Sci. 2007, 104, 11963-11968.

\bibitem{He2008}
He, Y.; Chen, Y.; Alexander, P.A.; Orban, J. NMR structures of two designed proteins with high sequence identity but different fold and function. Proc. Natl. Acad. Sci. 2008, 105, 14412-14417.

\bibitem{Alexander2009}
Alexander, P.A.; He, Y.; Chen, Y.; Orban, J. Bryan, P. A minimal sequence code for switching protein structure and function. Proc. Natl. Acad. Sci. 2009
, 106, 21149-21154.

\bibitem{Bryan2010}
Bryan, P.N.; Orban, J. Proteins that switch folds. Curr Opin Struct Biol. 2010, 20, 482-488.

\bibitem{He2012}
He, Y.; Chen, Y.; Alexander, P.A.; Bryan, P.N.; Orban, J. Mutational tipping points for switching protein folds and functions. Structure. 2012, 20,
83-91.

\bibitem{Sheffler2009}
Sheffler, W.; Baker, D. RosettaHoles: Rapid assessment of protein core packing for structure prediction, refinement, design, and validation. Protein Sci
ence. 2009, 18(1), 229-239.

\bibitem{Zhou2002}
Zhou H, Zhou Y. Distance-scaled, finite ideal-gas reference state improves structure-derived potentials of mean force for structure selection and
stability prediction. Protein Sci. 2002, 11, 2714-2726.

\bibitem{Allison2011}
Allison, J.R.; Bergeler, M.; van Gunsteren, W.F. Can computer modeling explain why two highly similar sequences fold into different structures ?
Biochemistry. 2011, 50, 10965-10973.

\bibitem{Handl2009}
Handl, J.; Knowles, J.; Lovel, S.C. Artefacts and biases affecting the evaluation of scoring
functions on decoy sets for protein structure prediction. Bioinformatics, 2009, 25, 1271-1279.

\bibitem{Kryshtafovych2011}
Kryshtafovych, A.; Fidelis, K; and Tramontano, A. Evaluation of model quality predictions in CASP9. Proteins, 2011, 79, 91–106

\bibitem{Wang2011}
Wang, Q.; Vantasin, K.; Xu, D.; Shang, Y. MUFOLD-WQA: A new selective consensus method for quality assessment in protein structure prediction.
Proteins, 2011, 79: 185-195.

\bibitem{Zemla2003}
Zemla, A. LGA: a method for finding 3D similarities in protein structures. Nucleic Acids Res 2003, 31, 3370–3374.

\bibitem{Case2012}
D.A. Case, T.A. Darden, T.E. Cheatham, III, C.L. Simmerling, J. Wang, R.E. Duke, R. Luo, R.C. Walker, W. Zhang, K.M. Merz, B. Roberts, S. Hayik, A. Roitberg, G. Seabra, J. Swails, A.W. Goetz, I. Kolossváry, K.F. Wong, F. Paesani, J. Vanicek, R.M. Wolf, J. Liu, X. Wu, S.R. Brozell, T. Steinbrecher, H. Gohlke, Q. Cai, X. Ye, J. Wang, M.-J. Hsieh, G. Cui, D.R. Roe, D.H. Mathews, M.G. Seetin, R. Salomon-Ferrer, C. Sagui, V. Babin, T. Luchko, S. Gusarov, A. Kovalenko, and P.A. Kollman (2012), AMBER 12, University of California, San Francisco.

\bibitem{Goetz2012}
Goetz, A.W.; Williamson, M.J.; Xu, D.; Poole, D.; Le Grand, S.; Walker, R.C. Routine microsecond molecular dynamics simulations with AMBER - Part I: Generalized
Born. J. Chem. Theory Comput. 2012, 8(5) 1542.

\bibitem{Roe2007}
Roe, D.R.; Okur, A.; Wickstrom, L.; Hornak, V.; Simmerling, C. Secondary structure bias in generalized Born solvent models: comparison of conformational ensembles 
and free energy of solvent polarization from explicit and implicit solvation. J Phys Chem B. 2007, 111, 1846-1857.

\bibitem{Case2005}
Case, D.A.; Cheatham, III, T.E.; Darden, T.; Gohlke, Luo, H.R.; Merz, Jr., K.M.;  Onufriev, A; Simmerling, C.;
Wang, B.; R. Woods, R. The Amber biomolecular simulation programs. J. Computat. Chem. (20005) 26, 1668-1688.

\bibitem{Hornak2006}
Hornak, V.; Abel, R.; Okur, A.; Strockbine, B.; Roitberg, A.; Simmerling, C. Comparison of multiple Amber force fields
and development of improved protein backbone parameters. Proteins., 2006, 65, 712-725.

\bibitem{Mongan2006}
Mongan, J.; Simmerling, C.; A. McCammon, J.; A. Case, D.; Onufriev, A. Generalized
Born with a simple, robust molecular volume correction. J. Chem. Theory Comput.,
2006, 3, 156-169.

\end{thebibliography}

\end{document}